---
title: "X-Wide Association Study in HIV+ in Zambian Females"
---

```{r setup, include=FALSE}
## hide the above
## commit to GitHub

knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggrepel)
library(plotly)
library(DT)
load('../merged_output/merged.Rdata')

mergedResults$merged_data$consistent_07_13 <- mergedResults$merged_data$replicated_univariate_07 & mergedResults$merged_data$replicated_univariate_13
```

## Authors (and contact):
- Chirag J Patel (chirag[at]hms[dot]harvard[dot]edu)
- Jay Bhattacharya 
- John PA Ioannidis
- Eran Bendavid (ebd[at]stanford[dot]edu)

(Submitted, 2017)

## Univariate -log10(pvalue) vs. association sizes in 2007 and 2013-2014
- Consistent (replicated in 2007 and 2013-2014 are shown in cyan)

```{r, echo=FALSE}
labelVars <- subset(mergedResults$merged_data, consistent_07_13 == T)
first <- mergedResults$merged_data[, c('consistent_07_13', 'variable', 'var_label', 'value_label', 'estimate_univariate_07', 'pvalue_univariate_07')]
first$year <- '2007'
second <- mergedResults$merged_data[, c('consistent_07_13', 'variable', 'var_label', 'value_label', 'estimate_univariate_13', 'pvalue_univariate_13')]
second$year <- '2013-2014'
colnames(first) <- colnames(second) <- c('consistent_07_13', 'variable', 'var_label', 'value_label', 'estimate', 'pvalue', 'year')
toPlot <- rbind(first, second)

toPlot$plot_label <- toPlot$var_label
toPlot$plot_label[toPlot$value_label != ''] <- paste(toPlot$var_label[toPlot$value_label != ''], 
                                                         '(', toPlot$value_label[toPlot$value_label != ''], ')', sep="")
```


```{r, echo=FALSE, width=7, height=5}
toLabel <- subset(toPlot, consistent_07_13==TRUE)
labelVarsTop <- labelVars[order(labelVars$pvalue_univariate_13, decreasing=F)[1:10], ]
toLabel <- toLabel[toLabel$variable %in% labelVarsTop$variable, ]
#toPlot$odds_ratio <- exp(toPlot$estimate)
p <- ggplot(toPlot, aes(exp(estimate), -log10(pvalue), 
                        color=consistent_07_13,
                        text=plot_label))
p <- p + geom_point(alpha=.5) + scale_x_log10() + facet_grid(. ~ year)
p <- p + theme(legend.position="none") + xlab('Odds Ratio')
p <- ggplotly(p)
p
```

### Zooming in on the consistent results only
```{r, echo=FALSE, width=7, height=5}
p <- ggplot(toLabel, aes(exp(estimate), -log10(pvalue)))
p <- p + geom_point() + geom_text_repel(data=toLabel, mapping=aes(exp(estimate), -log10(pvalue),label=plot_label), size=2, color='black')
p <- p + facet_grid(. ~ year)
p <- p + theme(legend.position="none") + xlab('Odds Ratio')
p
```

## Univariate models in 2007 and 2013-2014

- OR_07: Odds ratio in 2007 (nlogp_07 is the -log10(pvalue) of association)
- OR_13: Odds ratio in 2013 (nlogp_13 is the -log10(pvalue) of association)
- consistent_07_13: Finding that was replicated in the univariate model in both 2007 and 2013-2014
- For binary variables, the odds ratio is interpreted for the group denoted in the 'value_label' column.


```{r, echo=FALSE}
toTable <- mergedResults$merged_data[, c('consistent_07_13', 'variable_univariate_07', 'var_label', 'value_label',                                     'odds_ratio_univariate_07','odds_ratio_univariate_13', 
'pvalue_univariate_07', 'pvalue_univariate_13')]

toTable$OR_07 <- round(toTable$odds_ratio_univariate_07,2)
toTable$OR_13 <- round(toTable$odds_ratio_univariate_13,2)
toTable$odds_ratio_univariate_07 <- NULL
toTable$odds_ratio_univariate_13 <- NULL

toTable$nlogp_07 <- round(-log10(toTable$pvalue_univariate_07),2)
toTable$nlogp_13 <- round(-log10(toTable$pvalue_univariate_13),2)
toTable$pvalue_univariate_07 <- NULL
toTable$pvalue_univariate_13 <- NULL
toTable$varname <- toTable$variable_univariate_07
toTable$variable_univariate_07 <- NULL
toTable <- toTable[, c('varname', 'consistent_07_13', 'var_label', 'value_label', 'OR_07','OR_13', 
'nlogp_07', 'nlogp_13')]

DT::datatable(toTable, rownames=F)
```


## Associations under different modeling scenarios

* Univariate: no adjustments in association with HIV+ status
* *Ex ante* or "apriori"" model: 
    + v190_1, 2, 3, 4, 5: wealth index (1, 2, 3, 4, 5)
    + v025: type of place of residence (0,1)
    + v525_1: recent sexual activity (0,1)
    + v010: repondent's year of birth (as year)
    + v190_5: ever tested for HIV

  
```{r, echo=FALSE}
datModel07 <- mergedResults$merged_data[, c('consistent_07_13', 'variable', 'var_label', 'value_label', 'estimate_univariate_07', 'estimate_apriori_adj_07')]
datModel07$plot_label <- datModel07$var_label
datModel07$plot_label[datModel07$value_label != ''] <- paste(datModel07$var_label[datModel07$value_label != ''], 
                                                         '(', datModel07$value_label[datModel07$value_label != ''], ')', sep="")

p <- ggplot(datModel07, aes(estimate_univariate_07, estimate_apriori_adj_07, color=consistent_07_13, text=plot_label))
p <- p + geom_point() + geom_abline() 
ggplotly(p)

## 2013-2014
datModel13 <- mergedResults$merged_data[, c('consistent_07_13', 'variable', 'var_label', 'value_label', 'estimate_univariate_13', 'estimate_aprior_adj_13')]

datModel13$plot_label <- datModel13$var_label
datModel13$plot_label[datModel13$value_label != ''] <- paste(datModel13$var_label[datModel13$value_label != ''], 
                                                         '(', datModel13$value_label[datModel13$value_label != ''], ')', sep="")

p <- ggplot(datModel13, aes(estimate_univariate_13, estimate_aprior_adj_13, color=consistent_07_13, text=plot_label))
p <- p + geom_point() + geom_abline() 
ggplotly(p)
```